---
title: "ST558: Homework 4"
authors: "Scott Van Slyck"
description: "Advanced functions in R and querying an API"
date: "June 25, 2024"
format: html
editor: visual
---

```{r} 
#| echo: FALSE

library(tidyverse)
```

## Task 1: Conceptual Questions
1. What is the purpose of the lapply() function? What is the equivalent purrr function?


The lapply function applies a function to each element of a list and returns a list, the 'purrr' equivalent for this function is map().

2. Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns
are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall")
on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify
method = "kendall" when calling lapply())

lapply(my_list, FUN = cor, method = "kendall")


3. What are two advantages of using purrr functions instead of the BaseR apply family?

Firstly, 'purrr' functions like 'map()' provide consistent and easier to read syntax that integrates well with the rest of the tidyverse allowing code to be easier to understand. Secondly, 'purrr' offers type-specific functions like 'map_dbl()' and others to ensure the output is the desired type.

4. What is a side-effect function?

A side-effect function is a function that performs actions other than returning a value such as printing output, writing to a file, or interacting with an external system. They are called side-effects because they affect the program's state other than just producing a value or variable.


5. Why can you name a variable sd in a function and not cause any issues with the sd function?

Naming a variable 'sd' inside a function in R is safe because it remains local to the function and has a distinction different than the global sd() function. This variable would only run while the function is running.


## Task 2: Writing R functions
1. getRMSE function

```{r}
getRMSE <- function(response, prediction, ...){
  RMSE = sqrt(mean((response - prediction)^2, ...))
  return(RMSE = RMSE)
}

```

2. Testing RMSE function
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getRMSE(resp, pred)

resp[c(1,2)] = NA_real_
getRMSE(resp, pred)

getRMSE(resp, pred, na.rm = TRUE)

```

3/4. MAE function

```{r}
getMAE <- function(response, prediction, ...){
  MAE = mean(abs(response - prediction), ...)
  return(MAE = MAE)
}

set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getMAE(resp, pred)

resp[c(1,2)] = NA_real_
getMAE(resp, pred)

getMAE(resp, pred, na.rm = TRUE)

```

5/6. Wrapper function/Test
```{r}
metrics <- function(response, prediction, metrics = c("RMSE", "MAE"), ...){
  if(!(is.vector(response) & is.atomic(response) & is.numeric(response))){
  stop("response vector is not numeric.")
  }
  if(!(is.vector(prediction) & is.atomic(prediction) & is.numeric(prediction))){
  stop("prediction vector is not numeric.")
}

  results = list()
  
  if("RMSE" %in% metrics){
    results$RMSE = getRMSE(response, prediction)
  }
  
  if("MAE" %in% metrics){
    results$MAE = getMAE(response, prediction)
  }
  
  return(results)
}

set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))


metrics(resp, pred)

metrics(resp, pred, metrics = "RMSE")

metrics(resp, pred, metrics = "MAE")


resp[c(1,2)] = NA_real_
metrics(resp, pred)

metrics(resp, pred, metrics = "RMSE")

metrics(resp, pred, metrics = "MAE")


metrics(resp, pred, na.rm = TRUE)

metrics(resp, pred, metrics = "RMSE", na.rm = TRUE)

metrics(resp, pred, metrics = "MAE", na.rm = TRUE)

# metrics(resp, iris) presents error and does not allow me to run quarto render

```

## Task 3: API and Tidy-style function

1. news API reading
```{r}
library(httr)
dat = GET("https://newsapi.org/v2/top-headlines?sources=techcrunch&apiKey=dd989a70d3e549a08e899b8bee8c3707")
```


2. Parsing and finding data frame
```{r}
library(jsonlite)

parsed = dat$content |>
  rawToChar() |>
  fromJSON()
str(parsed, max.level = 1)

parsed$articles |>
  as_tibble()
```

3. Function to query the API
```{r}
news_pull <- function(title, from_date, api_key){
  url = "https://newsapi.org/v2/everything"
  params = list(
    q = title,
    from = from_date,
    sortBy = "publishedAt",
    apiKey = api_key
  )
  
  response = GET(url, query = params)
  
  content = content(response, "text")
  json_content = fromJSON(content, flatten = TRUE)
  
  return(json_content)
}
```